---
title: "Loading and Cleaning Sales Data"
author: "Kaushik Mohan"
date: "11/6/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## Data Munging packages
library(tidyverse)
library(dplyr)
library(lubridate)
library(readxl)
library(stringr)
library(utils)
library(Hmisc)
```

Looking at NYC Annualized Real-Estate sales data from 2012 onwards. All the files downloaded from https://www1.nyc.gov/site/finance/taxes/property-annualized-sales-update.page

## Loading Data

```{r}
boroughs <- c("manhattan","bronx","brooklyn","queens","statenisland")
years <- c("2017","2016","2015","2014","2013","2012")

boro_year <- expand.grid(boroughs,years)
colnames(boro_year) <- c("borough","year")
boro_year$file <- paste0(boro_year$year,"_",boro_year$borough,".xls")
```

```{r, warning=FALSE}
full_sales_data <- data.frame(matrix(NA,1,23))
error_cases <- NULL


for(i in c(1:dim(boro_year)[1])){
  
  tryCatch(
    {
      tmp <- read_excel(paste0("./data/raw/annualized_sales/",boro_year$file[i]),skip=4)
      ## standardizing column names to lowercase and removing spaces and escape characters
      colnames(tmp) <- gsub(" ","_",gsub("\n|\r","",tolower(colnames(tmp)))) 
      
      ## adding building class
      tmp$building_class <- substr(tmp$building_class_at_time_of_sale,1,1)
      
      ## Subsetting classes A,B,C,D,R,S as these are the 
      ## only ones coming under residential tax class
      tmp <- tmp %>% filter(building_class %in% c("A","B","C","D","R","S"))
      
      ## adding year
      tmp$year <- year(ymd(tmp$sale_date))
      
      colnames(full_sales_data) <- colnames(tmp)
      full_sales_data <- rbind(full_sales_data,tmp)
    }, 
    error=function(cond){
      error_cases <<- append(error_cases,paste0("../data/raw/annualized_sales/",boro_year$file[i]))
      print(paste0("../data/raw/annualized_sales/",boro_year$file[i]))
    }
  )
}

full_sales_data <- full_sales_data[-1,] ## removing first row of NAs

```

We note 9 cases where there is an error reading in the file. Looks like it's an issue with the size (correspondingly number of rows) and how .xls files deal with it. Saving them as `.xlsx` and reading them in seems to do the trick. 

```{r, warning=FALSE}
error_cases <- gsub(".xls",".xlsx",error_cases)

for(i in c(1:length(error_cases))){

    tmp <- read_excel(error_cases[i],skip=4)
    ## standardizing column names to lowercase and removing spaces and escape characters
    colnames(tmp) <- gsub(" ","_",gsub("\n|\r","",tolower(colnames(tmp)))) 
    
    ## adding building class
    tmp$building_class <- substr(tmp$building_class_at_time_of_sale,1,1)
    
    ## Subsetting classes A,B,C,D,R,S as these are the 
    ## only ones coming under residential tax class
    tmp <- tmp %>% filter(building_class %in% c("A","B","C","D","R","S"))
    
    ## adding year
    tmp$year <- year(ymd(tmp$sale_date))
    
    colnames(full_sales_data) <- colnames(tmp)
    full_sales_data <- rbind(full_sales_data,tmp)
}


```


## Merging with MapPluto Data

MapPLUTO Data has more detailed information about each building including their exact geospatial location. 

```{r}
load("../data/bbl_coords.RData")
merged_sales_data <- merge(full_sales_data,bbl_coords,by.x=c("borough","block","lot"),by.y=c("boro","Block","Lot"))

unmerged_sales_data <- anti_join(full_sales_data,bbl_coords,by=c("borough"="boro","block"="Block","lot"="Lot"))

round(dim(unmerged_sales_data)[1]*100/dim(full_sales_data)[1],2)
round(table(unmerged_sales_data$building_class)*100/table(full_sales_data$building_class),2)
round(table(unmerged_sales_data$building_class)*100/dim(full_sales_data)[1],2)
```

We note that we have several cases not merged properly with the MapPluto data. Closer inspection reveals that majority of these are building class R which are Condos. We'll deal with them separately. 

We use the GeoClient API to get the locations for the Condos. Details in script `03_geoclientAPI_data.R`. 

```{r}
## loading data merged with GeoClient API
load("../data/intermediary/missing_sales_data.RData")
```



```{r}

missing_sales_data$BBL <- as.numeric(missing_sales_data$BBL)
missing_sales_data <- merge(missing_sales_data,bbl_coords,by="BBL") 

# sales_data <- 
missing_sales_data <- missing_sales_data[,names(merged_sales_data)]


unmerged_missing_sales_data <- anti_join(missing_sales_data,bbl_coords,by="BBL") 

# sum(table(unmerged_missing_sales_data$building_class))/dim(unmerged_sales_data)[1]
```


```{r}
sales_data <- rbind(merged_sales_data,missing_sales_data)
```


## Merging with Property Assessment Roll Data

Lastly, we merge Property Assessment Roll archives data for more accurate information on apartment square footage. This is particularly essential as it corrects for condo unit square footage and also reduces the amount of missing data. We use the latest version of the assessment data FY18 for all the tax classes. The link to the data is below:
https://www1.nyc.gov/site/finance/taxes/property-assessment-roll-archives.page 

```{r}
assessment_tc1 <- mdb.get("../data/raw/assessment/tc1.mdb",tables="tc1")
assessment_tc234 <- mdb.get("../data/raw/assessment/TC234.mdb",tables="tc234")
assessment_data <- rbind(assessment_tc1,assessment_tc234)

cols_needed <- c("BBLE","BORO","BLOCK","LOT","EASE","TOT.UNIT","RES.UNIT","LND.AREA","GR.SQFT",
"CONDO.NM","CONDO.S1","CONDO.S2","CONDO.S3","CONDO.A","COMINT.L","COMINT.B",
"APTNO","AP.BORO","AP.BLOCK","AP.LOT","AP.EASE","GEO.RC","COOP.NUM")

assessment_data <- assessment_data[,cols_needed]
colnames(assessment_data)[1] <- "BBL"

# assessment_data$new_BBL <- as.numeric(paste0(assessment_data$BORO,
#                                   str_pad(assessment_data$BLOCK,5,side="left",pad="0"),
#                                   str_pad(assessment_data$LOT,4,side="left",pad="0")))

## removing easements
assessment_data <- assessment_data[assessment_data$EASE == "",]

```


```{r, warning=FALSE}
  
unmerged_sales_data <- anti_join(sales_data,assessment_data,by=c("borough"="BORO","block"="BLOCK","lot"="LOT"))

round(dim(unmerged_sales_data)[1]*100/dim(full_sales_data)[1],2)
round(table(unmerged_sales_data$building_class)*100/table(full_sales_data$building_class),2)
round(table(unmerged_sales_data$building_class)*100/dim(full_sales_data)[1],2)

sales_data <- merge(sales_data,assessment_data,
                                 by.x=c("borough","block","lot"),
                                 by.y=c("BORO","BLOCK","LOT"))

```


```{r}
# save(sales_data,file="../data/pre_cleaning_sales_data.RData")
```

