---
title: "Clusering Visualization"
author: "Kaushik Mohan"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = FALSE)
```


```{r}
load("../data/sales_features_2017.RData")
sales_features <- sales_features[match(census_tracts@data$boro_ct201,sales_features$boro_ct201),]
```

```{r}
#### Price Map
shades = auto.shading(sales_features$avg_price_1y)
choropleth(census_tracts,sales_features$avg_price_1y)
choro.legend(-74.455,40.93,shades,fmt="$%4.0f",cex=0.8,title='Avg. price per sq. ft')

shades = auto.shading(sales_features$avg_price_3y)
choropleth(census_tracts,sales_features$avg_price_3y)
choro.legend(-74.455,40.93,shades,fmt="$%4.0f",cex=0.75,
             title='3y moving average \n price per sq. ft')

shades = auto.shading(sales_features$avg_price_5y)
choropleth(census_tracts,sales_features$avg_price_5y)
choro.legend(-74.455,40.93,shades,fmt="$%4.0f",cex=0.75,
             title='5y moving average \n price per sq. ft')

```

# Clustering

```{r}
tracts_to_exclude <- c(census_pop$boro_ct201[census_pop$pop_2010 <= 500],"1023802") ##Rosevelt Island

reduced_tracts <- census_tracts[!(census_tracts$boro_ct201 %in% tracts_to_exclude),]

sales_features <- sales_features[!(sales_features$boro_ct201 %in% tracts_to_exclude),]
### geographical clustering
features_to_use <- !grepl("year|lon|lat|boro_ct201",colnames(sales_features))
features_to_use <- grepl("avg|sd|age|res",colnames(sales_features))

D0 <- dist(scale(sales_features[,features_to_use]))

list.nb <- poly2nb(reduced_tracts)
A <- nb2mat(list.nb,style = "B",zero.policy = TRUE)
diag(A) <- 1
D1 <- as.dist(1-A)
```


## Reshaped Boroughs

5 cluster solution

```{r}
K <- 5
colours <- rainbow(K)
tree <- hclustgeo(D0,D1,alpha=0.2)
clusters <- data.frame(cl=cutree(tree,K))
clusters$boro_ct201 <- sales_features$boro_ct201
clusters <- left_join(data.frame(boro_ct201=as.character(census_tracts$boro_ct201)),
                      clusters,by="boro_ct201")
clusters$cl[is.na(clusters$cl)] <- K+1
colours <- c(colours,"#A9A9A9A9") ## grey for excluded tracts
plot(census_tracts,col=colours[clusters$cl])

#table(clusters$cl)

```


## Reshaped Neighborhoods

55 cluster solution. 55 is the number of Public Use Microdata Areas (PUMA).

```{r}
K <- 55
colours <- distinctColorPalette(K)
tree <- hclustgeo(D0,D1,alpha=0.1)
clusters <- data.frame(cl=cutree(tree,K))
clusters$boro_ct201 <- sales_features$boro_ct201
clusters <- left_join(data.frame(boro_ct201=as.character(census_tracts$boro_ct201)),
                      clusters,by="boro_ct201")
clusters$cl[is.na(clusters$cl)] <- K+1
colours <- c(colours,"#A9A9A9A9") ## grey for excluded tracts
plot(census_tracts,col=colours[clusters$cl])

```

## Optimal solution 

Need to evaluate optimal number of clusters somehow. Around 100 seems reasonable?

```{r}
K <- 100
colours <- distinctColorPalette(K)
colours <- rainbow(K)
tree <- hclustgeo(D0,D1,alpha=0.1)
clusters <- data.frame(cl=cutree(tree,K))
clusters$boro_ct201 <- sales_features$boro_ct201
clusters <- left_join(data.frame(boro_ct201=as.character(census_tracts$boro_ct201)),
                      clusters,by="boro_ct201")
clusters$cl[is.na(clusters$cl)] <- K+1
colours <- c(colours,"#A9A9A9A9") ## grey for excluded tracts
plot(census_tracts,col=colours[clusters$cl])

```

Ideas:
- gganimate 5-100 clusters

To Do:
- shiny app
- add facilities db
- make map bigger
- showing features used
- new neighborhoood (without tract boundaries) on current PUMA map

Upload -> Merge -> select columns -> choose number of clusters -> run -> visualize





